---
title: 'Demo'
output: pdf_document
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


Some example calls of NH3/CO2/slurry pH model.

Load all functions
```{r}
source('../R/NH3_mods.R')
```

Use eqSpec to calculate equilibrium speciation
```{r}
eq <- eqSpec(tot = c(H. = -0.02, NH3 = 0.1, H2CO3 = 0.1, K. = 0.03, 
                     Na. = 0.02, Cl. = 0.03, HAc = 0.0242), 
             temp.c = 20, of = 'all')
eq
```

Some info on arguments: tot = molalities (concentration in mol/kg water) of "components" (elements here, but not below, each represented by a "master species"), temp.c = temperature in C, of = output format ('all' is the most complete).

Here is some info on the ouput from eqSpec
```{r}
names(eq)
```

m = molalities of species (conc. in mol/kg water), a = activities of species, g = activity coefficients, i = ionic strength, l.m = log molalities, l.a = log activities, tot = molalities of components (e.g., H2CO3 = TIC, NH3 = TAN), totk = similar to tot but here H2CO3 (i.e., H2CO3, HCO3-, and CO3-2 species) and CO2 (only CO2 (aq)) are listed separately--totk is only used to get totals for kinetic model where there are two IC components, cb = charge (un)balance, i.its = iterations used for ionic strength to stabilize, p.CO2 = equilibrium CO2 partial pressure
```{r}
str(eq)
eq
```

If you specify a pH, KOH or HAc (acetic acid) will be added to match it, so this call is essentially identical
```{r}
eq <- eqSpec(tot = c(H. = -0.02, NH3 = 0.1, H2CO3 = 0.1, 
                     K. = 0.03, Na. = 0.02, Cl. = 0.03, 
                     HAc = 0.0), 
             temp.c = 20, pH = 7.5, of = 'all')
```

Here we will extract total component concentrations for use in kinetic model (only difference from tot argument in call above is partitioning of H2CO3)
```{r}
itot <- eq$totk
```

itot is initial total concentration of components (including distribution of IC between H2CO3 and CO2, which are not necessarily in equilibrium in simulation below)
```{r}
itot
```

Now for the emission model (note that there is also an equilibrium version eqEmisDiffMod())
Next call predicts changes in solution chemistry as CO2 and NH3 are emitted from surface, assuming no carbonic anhydrase (that what the "kin" is for--kinetic)

May take 30 sec to run
```{r}
pred <- kinEmisDiffMod(c.thk = c(rep(1E-5, 5), rep(1E-4, 5), rep(9.45E-4, 10)), 
                       h.m = 1E-3, p.CO2.a = 4E-6, 
                       times = c(0:5*600, 1:12*3600), temp.c = 20, tot = itot)
```

Some info on arguments: c.thk = a vector of cell thicknesses in m (more and smaller is more accurate and slower), h.m = convective mass transfer coefficient in air for NH3 (m/s), p.CO2.a = ambient CO2 partial pressure (atm), temp.c = temperature in C, tot = molalities of components (usual length 8 vector and MUST be in order: "H.", "NH3", "H2CO3", "CO2", "K.", "Na.", "Cl.", "HAc")  

Info on the ouput
Complete output
```{r}
names(pred)
```
times = times, pos = cell positions (cell centers), tot.k = molalities of components that change with time (NH3, CO2, H2CO3) and note that msp stands for "master species", tot.f = molalities of compnents that are fixed (all other components are not transported, because there is no emission, and this model does not include "multi-component diffusion"), act = activities of each species, ph = pH at all locations and times, kin.CO2 = CO2 (aq) hydration rate (I think...need to check), emis = emission rates and cumulatives (a data frame).
str(pred)

Extact times
```{r}
tt <- pred$times
```
position (center of cells) converted to mm
```{r}
z <- pred$pos*1000
```
and pH
```{r}
ph <- pred$ph
```

Plot pH
```{r}
matplot(t(ph), z, type = 'l', ylim = c(10, 0))
```

"Surface" pH over time
```{r}
plot(tt, ph[, 1], type = 'o')
```

Get fluxes
```{r}
pred$emis
```

Plot NH3 flux
```{r}
plot(j.NH3 ~ t, data = pred$emis)
```

and cumulative emission
```{r}
plot(e.NH3 ~ t, data = pred$emis)
```




